---
title: "license data"
author: "Ashley Trudeau"
date: "6/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, lubridate, here, lme4, MASS, DHARMa, MuMIn, texreg, ggeffects, maps, mapproj, RColorBrewer, readxl)

```

Read in license sales data
```{r}
allsheets<-function(filename, tibble=T){
  sheets<-readxl::excel_sheets(filename)
  x<-lapply(sheets, function(X) read_excel(filename, sheet=X))
  if(!tibble) x<-lapply(x, as.data.frame)
  names(x)<-sheets
  x
}

data<-allsheets(here::here("license data","Copy of 2020 Sales Patterns1.1.xlsx"))
```


```{r}
# remove unexplained data in left columns
res.annual<-data$`RESIDENT ALL ANNUAL`[,-c(1:4)]
# rename columns
names(res.annual)<-as.character(res.annual[1,])
# drop first row (now names)
res.annual<-res.annual[-1,]
# fix column names to 

names(res.annual)<-c("monthDay","daily.2015","daily.2016","daily.2017","daily.2018","daily.2019","daily.2020",
                     "daily.avg.5.year","cum.2015","cum.2016","cum.2017","cum.2018","cum.2019","cum.2020","cum.avg.5.year","ref.cum.2019","ref.cum.5.year")


res.annual$sd.5year<-apply(res.annual[,c(9:13)], 1, sd)

sd.5year<-data.frame(monthDay=res.annual$monthDay, sd.5year=res.annual$sd.5year)
sd.5year.c<-sd.5year[complete.cases(sd.5year),]

counts.only<- dplyr::select(res.annual, monthDay,daily.2015,daily.2016,daily.2017,daily.2018,daily.2019,daily.2020,
                     daily.avg.5.year)
counts.only$daily.avg.5.year<-as.numeric(counts.only$daily.avg.5.year)

long.res.annual<-counts.only%>%
  pivot_longer(-monthDay, 
               names_to="year",
               values_to="daily.sales")

split<-str_split_fixed(long.res.annual$year, "daily.", 2)[,2]

long.res.annual$year<-str_split_fixed(long.res.annual$year, "daily.", 2)[,2]
# get rid of extra rows
long.res.annual<-filter(long.res.annual, !is.na(monthDay))

# split into average and annual dfs
resident.annual<-filter(long.res.annual, year!="avg.5.year")
resident.annual.mean<-filter(long.res.annual, year=="avg.5.year")

# get day of year for plots
resident.annual$date<-mdy(paste(resident.annual$monthDay, resident.annual$year, sep="/"))
resident.annual$doy<-yday(resident.annual$date)

# get cumulative sales
res.annual.sum<-resident.annual%>%
  group_by(year)%>%
  arrange(date)%>%
  mutate(cumulative.sales=cumsum(daily.sales))


# repeat day of year and cumulative sales for mean
resident.annual.mean$date<-mdy(paste(resident.annual.mean$monthDay, "2000", sep="/"))
resident.annual.mean$doy<-yday(resident.annual.mean$date)

res.annual.mean.sum<-resident.annual.mean%>%
  group_by(year)%>%
  arrange(date)%>%
  mutate(cumulative.sales=cumsum(daily.sales))



# get last three years
res.18.19<-filter(res.annual.sum, year%in%c("2018","2019","2020"))

# bind to 5 year average

res.18.19.5year<-rbind.data.frame(res.18.19, res.annual.mean.sum)

ggplot(res.18.19)+
  geom_line(aes(x=doy, y=cumulative.sales, color=year))+
    geom_line(data=res.annual.mean.sum, aes(x=doy, y=cumulative.sales), color="black")+
  theme_classic()


ggplot(res.annual.sum)+
  geom_line(aes(x=doy, y=cumulative.sales, color=year))+
  geom_line(data=res.annual.mean.sum, aes(x=doy, y=cumulative.sales), color="black")+
  theme_classic()



palette<-c( "#E69F00", "#56B4E9", "#CC79A7")



palette<-rev(palette)

# changing date format for plot

plot.date<-function(df){
  df$date.c<-as.character(df$date)
  split<-str_split_fixed(df$date.c,"-",3)
  df$plotDate<-paste(split[,2],split[,3],"2000", sep="-")
  df$plotDate<-mdy(df$plotDate)
  return(df)
}

res.plot<-plot.date(res.18.19.5year)
res.plot$year<-ifelse(res.plot$year=="avg.5.year","5 year average\n2015-2019",res.plot$year)
res.plot$Year<-res.plot$year

# trying to add error bars, so need to separate mean of 5 years from separate values

res.plot.mean<-filter(res.plot, year=="5 year average\n2015-2019")
res.plot.mean$sd<-sd.5year.c$sd.5year
# 
 res.plot.years<-filter(res.plot, year!="5 year average\n2015-2019")

resident<-ggplot()+
  geom_ribbon(data=res.plot.mean , aes(x=plotDate, ymin=cumulative.sales-1.96*sd, ymax=cumulative.sales+1.96*sd), fill="gray", alpha=0.75)+
  geom_line(data=res.plot.mean, aes(x=plotDate, y=cumulative.sales), color="black", lwd=1.5)+
  geom_line(data=res.plot.years, aes(x=plotDate, y=cumulative.sales, color=Year), lwd=1.5)+
  scale_x_date(labels = function(x) format(x, "%d-%b"), position="bottom")+ 
  scale_color_manual(values=palette)+
   ylab("Cumulative number of\nlicense sales")+
   xlab("Date")+
   theme_classic()+
  theme(legend.position="none")
resident
ggsave(here::here("figures","resident.sales.png"), height=4, width=6)



```


Nonresident annual
```{r}

# remove unexplained data in left columns
nonres.annual<-data$`NONRESIDENT ALL ANNUAL`[,-c(1:4)]
# rename columns
names(nonres.annual)<-as.character(nonres.annual[1,])
# drop first row (now names)
nonres.annual<-nonres.annual[-1,]
# fix column names to 

names(nonres.annual)<-c("monthDay","daily.2015","daily.2016","daily.2017","daily.2018","daily.2019","daily.2020",
                     "daily.avg.5.year","cum.2015","cum.2016","cum.2017","cum.2018","cum.2019","cum.2020","cum.avg.5.year","ref.cum.2019","ref.cum.5.year")
counts.only<- dplyr::select(nonres.annual, monthDay,daily.2015,daily.2016,daily.2017,daily.2018,daily.2019,daily.2020,
                     daily.avg.5.year)
counts.only$daily.avg.5.year<-as.numeric(counts.only$daily.avg.5.year)

long.nonres.annual<-counts.only%>%
  pivot_longer(-monthDay, 
               names_to="year",
               values_to="daily.sales")

split<-str_split_fixed(long.nonres.annual$year, "daily.", 2)[,2]

long.nonres.annual$year<-str_split_fixed(long.nonres.annual$year, "daily.", 2)[,2]
# get rid of extra rows
long.nonres.annual<-filter(long.nonres.annual, !is.na(monthDay))

# split into average and annual dfs
nonresident.annual<-filter(long.nonres.annual, year!="avg.5.year")
nonresident.annual.mean<-filter(long.nonres.annual, year=="avg.5.year")

# get day of year for plots
nonresident.annual$date<-mdy(paste(nonresident.annual$monthDay, nonresident.annual$year, sep="/"))
nonresident.annual$doy<-yday(nonresident.annual$date)

# get cumulative sales
nonres.annual.sum<-nonresident.annual%>%
  group_by(year)%>%
  arrange(date)%>%
  mutate(cumulative.sales=cumsum(daily.sales))


# repeat day of year and cumulative sales for mean
nonresident.annual.mean$date<-mdy(paste(nonresident.annual.mean$monthDay, "2000", sep="/"))
nonresident.annual.mean$doy<-yday(nonresident.annual.mean$date)

nonres.annual.mean.sum<-nonresident.annual.mean%>%
  group_by(year)%>%
  arrange(date)%>%
  mutate(cumulative.sales=cumsum(daily.sales))



# get last three years
nonres.18.19<-filter(nonres.annual.sum, year%in%c("2018","2019","2020"))


comp.dat.nonres<-rbind.data.frame(nonres.18.19, nonres.annual.mean.sum)


ggplot(nonres.18.19)+
  geom_line(aes(x=doy, y=cumulative.sales, color=year))+
  geom_line(data=nonres.annual.mean.sum, aes(x=doy, y=cumulative.sales), color="black")+
  theme_classic()


ggplot(nonres.annual.sum)+
  geom_line(aes(x=doy, y=cumulative.sales, color=year))+
  geom_line(data=nonres.annual.mean.sum, aes(x=doy, y=cumulative.sales), color="black")+
  theme_classic()


ggplot(comp.dat.nonres)+
  geom_line(aes(x=doy, y=cumulative.sales, color=year))+
  theme_classic()


# changing date format for plot

nonres.18.19.5year<-comp.dat.nonres



nonres.plot<-plot.date(nonres.18.19.5year)
nonres.plot$year<-ifelse(nonres.plot$year=="avg.5.year","5 year average\n2015-2019",nonres.plot$year)
nonres.plot$Year<-nonres.plot$year


nonres.annual$sd.5year<-apply(nonres.annual[,c(9:13)], 1, sd)

sd.5year<-data.frame(monthDay=nonres.annual$monthDay, sd.5year=nonres.annual$sd.5year)
sd.5year.c<-sd.5year[complete.cases(sd.5year),]


nonres.plot.mean<-filter(nonres.plot, year=="5 year average\n2015-2019")
nonres.plot.mean$sd<-sd.5year.c$sd.5year

 nonres.plot.years<-filter(nonres.plot, year!="5 year average\n2015-2019")



nonresident<-ggplot(nonres.plot)+
  geom_ribbon(data=nonres.plot.mean , aes(x=plotDate, ymin=cumulative.sales-2*sd, ymax=cumulative.sales+2*sd), fill="gray", alpha=0.75)+
  geom_line(data=nonres.plot.mean, aes(x=plotDate, y=cumulative.sales), color="black", lwd=1.5)+
  geom_line(data=nonres.plot.years, aes(x=plotDate, y=cumulative.sales, color=Year), lwd=1.5)+
    scale_x_date(labels = function(x) format(x, "%d-%b"), position="bottom") +
  scale_color_manual(values=palette)+
  ylab("Cumulative number of\nlicense sales")+
  xlab("Date")+
  theme_classic()+
  theme(legend.position="none")
nonresident
ggsave(here::here("figures","nonresident.sales.png"), height=4, width=6)




```

still higher than average, but not highest in past 5 years. 2015 saw more nonresident annual license sales. 


Now resident and nonresident first time buyers

```{r}
# remove unexplained data in left columns
res.ftb<-data$`RESIDENT FTB`[,-c(1:5)]
# rename columns
names(res.ftb)<-as.character(res.ftb[1,])
# drop first row (now names)
res.ftb<-res.ftb[-1,]
# fix column names to 

names(res.ftb)<-c("monthDay","daily.2015","daily.2016","daily.2017","daily.2018","daily.2019","daily.2020",
                     "daily.avg.5.year","cum.2015","cum.2016","cum.2017","cum.2018","cum.2019","cum.2020","cum.avg.5.year","ref.cum.2019","ref.cum.5.year")
counts.only<- dplyr::select(res.ftb, monthDay,daily.2015,daily.2016,daily.2017,daily.2018,daily.2019,daily.2020,
                     daily.avg.5.year)
counts.only$daily.avg.5.year<-as.numeric(counts.only$daily.avg.5.year)

long.res.ftb<-counts.only%>%
  pivot_longer(-monthDay, 
               names_to="year",
               values_to="daily.sales")

split<-str_split_fixed(long.res.ftb$year, "daily.", 2)[,2]

long.res.ftb$year<-str_split_fixed(long.res.ftb$year, "daily.", 2)[,2]
# get rid of extra rows
long.res.ftb<-filter(long.res.ftb, !is.na(monthDay))

# split into average and ftb dfs
resident.ftb<-filter(long.res.ftb, year!="avg.5.year")
resident.ftb.mean<-filter(long.res.ftb, year=="avg.5.year")

# get day of year for plots
resident.ftb$date<-mdy(paste(resident.ftb$monthDay, resident.ftb$year, sep="/"))
resident.ftb$doy<-yday(resident.ftb$date)

# get cumulative sales
res.ftb.sum<-resident.ftb%>%
  group_by(year)%>%
  arrange(date)%>%
  mutate(cumulative.sales=cumsum(daily.sales))


# repeat day of year and cumulative sales for mean
resident.ftb.mean$date<-mdy(paste(resident.ftb.mean$monthDay, "2000", sep="/"))
resident.ftb.mean$doy<-yday(resident.ftb.mean$date)

res.ftb.mean.sum<-resident.ftb.mean%>%
  group_by(year)%>%
  arrange(date)%>%
  mutate(cumulative.sales=cumsum(daily.sales))



# get last three years
res.ftb.18.19<-filter(res.ftb.sum, year%in%c("2018","2019","2020"))

# bind on 5 year average
res.ftb.18.19.5year<-rbind.data.frame(res.ftb.18.19, res.ftb.mean.sum)

# fix the dates for the plot

res.ftb.plot<-plot.date(res.ftb.18.19.5year)
res.ftb.plot$year<-ifelse(res.ftb.plot$year=="avg.5.year","5 year average\n2015-2019",res.ftb.plot$year)
res.ftb.plot$Year<-res.ftb.plot$year

res.ftb$sd.5year<-apply(res.ftb[,c(9:13)], 1, sd)

sd.5year<-data.frame(monthDay=res.ftb$monthDay, sd.5year=res.ftb$sd.5year)
sd.5year.c<-sd.5year[complete.cases(sd.5year),]


res.ftb.plot.mean<-filter(res.ftb.plot, year=="5 year average\n2015-2019")
res.ftb.plot.mean$sd<-sd.5year.c$sd.5year

 res.ftb.plot.years<-filter(res.ftb.plot, year!="5 year average\n2015-2019")


resident.ftb<-ggplot(res.ftb.plot)+
  geom_ribbon(data=res.ftb.plot.mean , aes(x=plotDate, ymin=cumulative.sales-2*sd, ymax=cumulative.sales+2*sd), fill="gray", alpha=0.75)+
  geom_line(data=res.ftb.plot.mean, aes(x=plotDate, y=cumulative.sales), color="black", lwd=1.5)+
  geom_line(data=res.ftb.plot.years, aes(x=plotDate, y=cumulative.sales, color=Year), lwd=1.5)+
    scale_x_date(labels = function(x) format(x, "%d-%b"), position="bottom") +
  scale_color_manual(values=palette)+
  ylab("Cumulative number of\nlicense sales")+
  xlab("Date")+
  theme_classic()+
  theme(legend.position="none")
resident.ftb
ggsave(here::here("figures","resident.ftb.sales.png"), height=4, width=6)







```


Nonresident FTB

```{r}
# remove unexplained data in left columns
nonres.ftb<-data$`NONRESIDENT FTB`[,-c(1:5)]
# rename columns
names(nonres.ftb)<-as.character(nonres.ftb[1,])
# drop first row (now names)
nonres.ftb<-nonres.ftb[-1,]
# fix column names to 

names(nonres.ftb)<-c("monthDay","daily.2015","daily.2016","daily.2017","daily.2018","daily.2019","daily.2020",
                     "daily.avg.5.year","cum.2015","cum.2016","cum.2017","cum.2018","cum.2019","cum.2020","cum.avg.5.year","ref.cum.2019","ref.cum.5.year")
counts.only<- dplyr::select(nonres.ftb, monthDay,daily.2015,daily.2016,daily.2017,daily.2018,daily.2019,daily.2020,
                     daily.avg.5.year)
counts.only$daily.avg.5.year<-as.numeric(counts.only$daily.avg.5.year)

long.nonres.ftb<-counts.only%>%
  pivot_longer(-monthDay, 
               names_to="year",
               values_to="daily.sales")

split<-str_split_fixed(long.nonres.ftb$year, "daily.", 2)[,2]

long.nonres.ftb$year<-str_split_fixed(long.nonres.ftb$year, "daily.", 2)[,2]
# get rid of extra rows
long.nonres.ftb<-filter(long.nonres.ftb, !is.na(monthDay))

# split into average and ftb dfs
nonresident.ftb<-filter(long.nonres.ftb, year!="avg.5.year")
nonresident.ftb.mean<-filter(long.nonres.ftb, year=="avg.5.year")

# get day of year for plots
nonresident.ftb$date<-mdy(paste(nonresident.ftb$monthDay, nonresident.ftb$year, sep="/"))
nonresident.ftb$doy<-yday(nonresident.ftb$date)

# get cumulative sales
nonres.ftb.sum<-nonresident.ftb%>%
  group_by(year)%>%
  arrange(date)%>%
  mutate(cumulative.sales=cumsum(daily.sales))


# repeat day of year and cumulative sales for mean
nonresident.ftb.mean$date<-mdy(paste(nonresident.ftb.mean$monthDay, "2000", sep="/"))
nonresident.ftb.mean$doy<-yday(nonresident.ftb.mean$date)

nonres.ftb.mean.sum<-nonresident.ftb.mean%>%
  group_by(year)%>%
  arrange(date)%>%
  mutate(cumulative.sales=cumsum(daily.sales))



# get last three years
nonres.ftb.18.19<-filter(nonres.ftb.sum, year%in%c("2018","2019","2020"))

# bind on 5 year average
nonres.ftb.18.19.5year<-rbind.data.frame(nonres.ftb.18.19, nonres.ftb.mean.sum)

# fix the dates for the plot

nonres.ftb.plot<-plot.date(nonres.ftb.18.19.5year)
nonres.ftb.plot$year<-ifelse(nonres.ftb.plot$year=="avg.5.year","5 year average\n2015-2019",nonres.ftb.plot$year)
nonres.ftb.plot$Year<-nonres.ftb.plot$year

nonres.ftb$sd.5year<-apply(nonres.ftb[,c(9:13)], 1, sd)

sd.5year<-data.frame(monthDay=nonres.ftb$monthDay, sd.5year=nonres.ftb$sd.5year)
sd.5year.c<-sd.5year[complete.cases(sd.5year),]


nonres.ftb.plot.mean<-filter(nonres.ftb.plot, year=="5 year average\n2015-2019")
nonres.ftb.plot.mean$sd<-sd.5year.c$sd.5year

 nonres.ftb.plot.years<-filter(nonres.ftb.plot, year!="5 year average\n2015-2019")



nonresident.ftb<-ggplot(nonres.ftb.plot)+
  geom_ribbon(data=nonres.ftb.plot.mean , aes(x=plotDate, ymin=cumulative.sales-2*sd, ymax=cumulative.sales+2*sd), fill="gray", alpha=0.75)+
  geom_line(data=nonres.ftb.plot.mean, aes(x=plotDate, y=cumulative.sales), color="black", lwd=1.5)+
  geom_line(data=nonres.ftb.plot.years, aes(x=plotDate, y=cumulative.sales, color=Year), lwd=1.5)+
    scale_x_date(labels = function(x) format(x, "%d-%b"), position="bottom") +
  scale_color_manual(values=palette)+
  ylab("Cumulative number of\nlicense sales")+
  xlab("Date")+
  theme_classic()+
  theme(legend.position="none")
nonresident.ftb
ggsave(here::here("figures","nonresident.ftb.sales.png"), height=4, width=6)




```

Big increase in nonresident FTB license sales (as opposed to more modest increase in regular annual). I think I remember most of the increase in nonres licenses driven by these FTBs?


```{r}
library(cowplot)

palette<-c("#000000", "#E69F00", "#56B4E9", "#CC79A7")



palette<-rev(palette)

legend.plot<-ggplot(nonres.ftb.plot)+
  geom_line(aes(x=plotDate, y=cumulative.sales, color=Year), lwd=1)+
    scale_x_date(labels = function(x) format(x, "%d-%b"), position="bottom") +
    ylim(0,90000)+
  scale_color_manual(values=palette)+
  ylab("Cumulative sales")+
  xlab("Date")+
  theme_classic()


legend<-get_legend(legend.plot + theme(legend.box.margin=margin(0,0,0,0)))

plot_grid(resident, nonresident, legend, resident.ftb, nonresident.ftb, labels=c("A","B","","C","D"),  nrow=2, ncol=3, rel_widths=c(1,1,0.4))
ggsave(here::here("figures","all.license.sales.png"), height=6, width=9)
```

What proportion of the increase in resident and nonresident sales in 2020 came from first time buyer licenses? 

```{r}
res.max<-res.plot%>%
  group_by(year)%>%
  summarize(total=max(cumulative.sales))

nonres.max<-nonres.plot%>%
  group_by(year)%>%
  summarize(total=max(cumulative.sales))

ftb.res<-res.ftb.plot%>%
  group_by(year)%>%
  summarize(total=max(cumulative.sales))

ftb.nonres<-nonres.ftb.plot%>%
  group_by(year)%>%
  summarize(total=max(cumulative.sales))

res.max
ftb.res

```
Change in resident licenses

```{r}
1019570-941940
```
change in res ftb licenses

```{r}
87730-51206
```
```{r}
36524/77630
```
47% of resident license sales increase came from FTB

nonresidents

```{r}
nonres.max
ftb.nonres
```

Change in nonres
```{r}
188034-174627
```
Change in ftb
```{r}
45088-33481
```
```{r}
11607/13407
```
87% of nonres license increase was FTBs


Can I plot license sales versus vehicle counts? 

```{r}
vehicles<-read_csv("model.data.csv")

```

```{r}
res.2018<-res.annual%>%
  dplyr::select( monthDay, daily.2018)%>%
  mutate(year=rep(2018),
         type="resident")%>%
  rename(daily.sold=daily.2018)

res.2019<-res.annual%>%
  dplyr::select(monthDay, daily.2019)%>%
  mutate(year=rep(2019),
         type="resident")%>%
  rename(daily.sold=daily.2019)

res.2020<-res.annual%>%
  dplyr::select(monthDay, daily.2020)%>%
  mutate(year=rep(2020),
         type="resident")%>%
  rename(daily.sold=daily.2020)


nonres.2018<-nonres.annual%>%
  dplyr::select( monthDay, daily.2018)%>%
  mutate(year=rep(2018),
         type="nonresident")%>%
  rename(daily.sold=daily.2018)

nonres.2019<-nonres.annual%>%
  dplyr::select(monthDay, daily.2019)%>%
  mutate(year=rep(2019),
         type="nonresident")%>%
  rename(daily.sold=daily.2019)

nonres.2020<-nonres.annual%>%
  dplyr::select(monthDay, daily.2020)%>%
  mutate(year=rep(2020),
         type="nonresident")%>%
  rename(daily.sold=daily.2020)

license<-rbind.data.frame(res.2018, res.2019, res.2020, nonres.2018, res.2019, nonres.2020)

license.all<-license%>%
  group_by(year, monthDay)%>%
  summarize(total.license=sum(daily.sold))%>%
  ungroup()%>%
  mutate(date=mdy(paste(monthDay, year, sep="/")))%>%
  mutate(dayOfYear=yday(date))

```

now join to vehicle count data

```{r}
#vehicles$dayOfYear<-yday(vehicles$date)

vehicles.license<-left_join(vehicles, license.all[,c("date","total.license","dayOfYear")], by="date")
vehicles.license$year<-as.factor(vehicles.license$year)

```
```{r}

palette<-c("#E69F00", "#56B4E9", "#CC79A7")



palette<-rev(palette)

license.all$year<-as.factor(license.all$year)

license.plot.no.leg<-ggplot(license.all)+
  geom_point(aes(x=dayOfYear, y=total.license, color=year))+
  scale_color_manual(values=palette)+
  xlab("Day of year")+
  xlim(50, 275)+
  ylab("Daily licenses sold")+
  theme_classic()+
  theme(legend.position="none")
```

```{r}
vehicles$year<-as.factor(vehicles$year)

vehicle.plot<-ggplot(vehicles)+
  geom_point(aes(x=dayOfYear, y=totalVehicles, color=year))+
  scale_color_manual(values=palette)+
  xlab("Day of year")+
  ylab("Vehicle count observations")+
  theme_classic()

legend<-get_legend(vehicle.plot)

vehicle.plot.no.leg<-ggplot(vehicles)+
  geom_point(aes(x=dayOfYear, y=totalVehicles, color=year))+
  scale_color_manual(values=palette)+
  xlab("Day of year")+
  ylab("Vehicle count observations")+
  xlim(50, 275)+
  theme_classic()+
  theme(legend.position="none")
```


```{r}
plot_grid(license.plot.no.leg, vehicle.plot.no.leg, legend, nrow=1, labels=c("A","B"," "), rel_widths = c(0.45, 0.45, 0.1))
ggsave(here::here("figures","seasonality.sales.vehicles.png"), height=4, width=8)
```



```{r}
ggplot(vehicles)+
  geom_boxplot(aes(x=year, y=totalVehicles))
```

```{r}
ggplot(vehicles)+
  geom_boxplot(aes(x=lakeID, y=totalVehicles, fill=year))+
  scale_fill_manual(values=palette)+
  ylab("Number of vehicles observed")+
  xlab("Lake ID")+
  theme_classic()+
  theme(text=element_text(size=11))
ggsave(here::here("figures","raw.boxplot.png"), height=4, width=10)
```

outliers removed--put them back? color code to show which ones

```{r}
ggplot(vehicles)+
  geom_point(aes(x=hourOfDay, y=totalVehicles, color=year))+
  scale_color_manual(values=palette)+
  theme_classic()+
  theme(text=element_text(size=15))
ggsave(here::here("figures","quad.time.png"), height=4, width=6)
```






